---
title: "Processing the SERO estimation output"
author: "Katy Gaythorpe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing the SERO estimation output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setupknit, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  warning = FALSE,
  message = FALSE
)
```

This guide is to help you examine the output of the product space MCMC for the SERO. We therefore assume that you already have a folder filled with Markov chain Monte Carlo samples from the posterior of the parameters for the seroprevalence models. 

```{r setup, fig.cap = "\\label{fig:trace}Trace plot of model indicator.", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8 }
library(mcmcplots)
library(ggmcmc)
library(gridExtra)
library(maptools)
library(readr)
library(fields)
library(Hmisc)

library(YFestimation)

snapal = "Fjord"

mcmc_out =get_chains(
  "../../multi_model_MCMC_chain_20180622", 
  burnin = 1e3, thin = 100)

#plot model chain
plot(mcmc_out$model_chain, type="l", ylim=c(0,1), 
     ylab="Model indicator")
```

The only shared parameters are the vaccine efficacy and vc factor for CMRs, the traceplots for these are as follows.

```{r trace,  message = F, warning=F, fig.cap = "\\label{fig:trace}Trace plot.", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8 }

mcmc_out2 = ggs( convert.mcmc.list( mcmc_out[,1:83]) )

f1 = ggs_traceplot(mcmc_out2, family = "vac")

f2 = ggs_traceplot(mcmc_out2, family = "vc")

grid.arrange(f1, f2, nrow=1)
```

```{r partial,  message = F, warning=F, fig.cap = "\\label{fig:partial}Density plot of entire chain (grey) and last 10\\% of the chain (green).", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8 }


f1 = ggs_compare_partial(mcmc_out2, family = "vac")+
  theme(legend.position = "none")

f2 = ggs_compare_partial(mcmc_out2, family = "vc")+ 
  theme(legend.position = "none")

grid.arrange(f1, f2, nrow=1)
```

```{r prior_posterior_vac_eff,  message = F, warning=F, fig.cap = "\\label{fig:priorposterior}Prior and posterior densities for vaccine efficacy and vc factor for CMRs.", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8}

MultiModelEstimation::plot_prior_post(mcmc_out, 
                                      snapalette::snapalette(snapal, 
                                                             3, 
                                                             type="discrete")[2], 
                                      "SERO")
```


\FloatBarrier 

## Bayes factors

We may calculate the Bayes factors by approximating the posterior model probability form the model-indicator chain. The proportion of time spent in either model is a proxy for the model evidence which we may use, with the model priors, to calculate the log Bayes factors. We calculate the model evidence using a boostrap sample of the model chain in order to capture some of the variation.


```{r bootstrap,  echo=F, message = F, warning=F, fig.cap = "\\label{fig:bootstrap}Bootrsap sample of model evidence for $R_0$.", fig.align='center', fig.pos = 'h', strip.white=F, fig.width=8 }

out = calculate_bootstrap_Bayes(mcmc_out, 
                                col = snapalette::snapalette(snapal, 
                                                             3, 
                                                             type="discrete")[2])
```

The log Bayes factor for $R_0$ over Foi is ` toString(format(round(out$log_Bayes_R0_FOI[2], 2), nsmall = 2))` in range [` toString(format(round(out$log_Bayes_R0_FOI[1], 2), nsmall = 2))`, ` toString(format(round(out$log_Bayes_R0_FOI[3], 2), nsmall = 2))`]. 


\FloatBarrier

##Assessing the fit

We will now plot the Hpd interval predictions of both transmission models.

```{r setup_fit,  message = F, warning=F}

# LOAD countries #
Countries = read_csv(paste0("../../../Data/","Countries.csv"))

# LOADING SEROLOGY DATA #
Serology = read.csv(paste0("../../../Data/","Serology/serology.csv"), stringsAsFactors = FALSE)
seroout = process_serology(Serology)

# LOAD envdat #
Env_Table_path = paste0("../../../Data/","Environment/Africa_adm1_dat_2017.csv") #this file is adapted by hand to have latest outbreaks
envdat = launch_env_dat(Env_Table_path, Countries$c34)

# POPULATION AND VACCINATION DATA #
all_res_pop_3d = get_pop_data_3d(path = "../../../Data/", 
                                 c_country = Countries$c34, 
                                 dat = envdat$dat) # can we get rid of envdat dependency?

pop1 = all_res_pop_3d$pop1                           
pop3d = all_res_pop_3d$pop3d                               
P_tot_2d = all_res_pop_3d$P_tot_2d                                   
p_prop_3d = all_res_pop_3d$p_prop_3d                                

#get names
dim_adm  = dimnames(pop3d)[[1]]
dim_year = as.numeric(dimnames(pop3d)[[2]])
dim_age  = dimnames(pop3d)[[3]]


# VACCINATION DATA #
vaccdir = paste0("../../../Data/", "Vaccination/")

latest_vaccine_csv = "Outputs/adm1_old/vaccination_coverage_by_adm1_year_age_base_skew0.csv"   #updated at end of 2017

vc2d = read.csv(paste0(vaccdir,latest_vaccine_csv), 
                stringsAsFactors = FALSE) #read in latest vaccine coverage estimates

names(vc2d)[names(vc2d)=="country"]= "adm0"                          #rename countries as adm0
names(vc2d)[names(vc2d)=="adm1"]= "adm0_adm1"                        #renames adm1 as adm0_adm1

# formally "repair_vc_data" from FOI model in Kevin's folder
for (colIndex in 3:ncol(vc2d)){                                      #before 1995, we have NA values for those aged >75
  vc2d[,colIndex] = ifelse(is.na(vc2d[,colIndex]), vc2d[,colIndex-1], vc2d[,colIndex])
}
# restrict to lines in dat
vc2d = vc2d[vc2d[,"adm0_adm1"] %in% envdat$dat[,"adm0_adm1"],]

#vc3d
vc3d = transform_into_vc3d(vc2d,  adm="adm1")

# t0_vac_africa #
t0_vac_africa = calc_t0_vac_africa(vc3d)

# inc_v3d #
inc_v3d = calc_incidence_vac_general(vc3d)
    
# CALCULATE population moments #
pop_moments_whole = calc_pop_moments(p_prop_3d, t0_vac_africa,dim_adm,dim_year,dim_age)

# aggregate #
list_aggregate_pop_vc = Make_aggregate_pop_vc_3d(pop1=pop1, vc2d=vc2d, sero_studies=seroout$sero_studies, adm1s=seroout$adm1s)
pop_agg3d = list_aggregate_pop_vc$pop_agg3d
vc_agg3d = list_aggregate_pop_vc$vc_agg3d

# calculate aggregated incidence (same function as before)
inc_v3d_agg = calc_incidence_vac_general(vc_agg3d);

# calculate aggregated moments (different fucntion before)
pop_moments_agg = calc_pop_moments_agg(pop_agg3d,seroout$t0_vac,dim_year,seroout$study_years)

# pop_momenets_whole #
pop_moments_whole = calc_pop_moments(p_prop_3d, 
                                     t0_vac_africa,
                                     dim_adm,
                                     dim_year,
                                     dim_age)
    

# CREATE R0 LOOKUP TABLE #
if(!file.exists(paste0("../../../YellowFeverModelEstimation2017/","R0_lookup_table.Rdata") )){
    
    ptm = proc.time()
    create_R0_lookup(dim_adm, 
                     envdat$dat, 
                     dim_year,
                     dim_age,
                     p_prop_3d,
                     P_tot_2d,
                     inc_v3d,
                     pop_moments_whole,
                     pop_moments_agg, 
                     vac_eff_arg = 0.975)
    proc.time()-ptm
  } else {
    load(paste0("../../../YellowFeverModelEstimation2017/","R0_lookup_table.Rdata") )
  }


# pop at survey #
foi_const_surv = rep(0, seroout$no_sero_surveys)

list_pop_at_survey = create_pop_at_survey(pop_agg3d, seroout$sero_studies, dim_year)
p_at_survey = list_pop_at_survey$p_at_survey_3d
P_tot_survey = list_pop_at_survey$P_tot_survey_2d


```


```{r hpd, message=F, warning=F}

out = calculate_hpd(mcmc_out, glm_mcmc_out = NA)

hpd_sero = out$hpd_sero

```

```{r sero_predict1,  message = F, warning=F, fig.cap = "\\label{fig:sero_predict1}Seroprevalence predictions from $R_0$ model (red) and Foi model (blue).", fig.align='center', fig.pos = 'h', fig.height = 10, fig.width = 20, strip.white=F}

plot_sero_predictions(seroout, hpd_sero, pop_agg3d, dim_year, 
                      dim_age, inc_v3d, pop_moments_agg, vc_agg3d)

```


