---
title: "Processing the GLM estimation output"
author: "Katy Gaythorpe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing the GLM estimation output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setupknit, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,
  warning = FALSE,
  message = FALSE
)
```

This guide is to help you examine the output of the MCMC for the GLM. We therefore assume that you already have a folder filled with Markov chain Monte Carlo samples from the posterior of the parameters for the generalised linear model.

```{r setup}
library(mcmcplots)
library(ggmcmc)
library(gridExtra)
library(maptools)
library(readr)
library(fields)

library(YFestimation)

glm_mcmc_out=get_chains("../../chains/GLM_MCMC_chain_20180515", 
                        burnin = 2e4, thin = 1)

glm_mcmc_out2 = ggmcmc::ggs( mcmcplots::convert.mcmc.list( 
                    subset( glm_mcmc_out, 
                            select = -c(posteriorProb, acceptRate))) )

snapal = "Venice"
```


To assess the convergence of the MCMC chains we shall plot a selection of the outputs for two parameters.

```{r trace_glm,   fig.cap = "\\label{fig:trace_glm}Trace plot.", fig.align='center', fig.pos = 'h', strip.white=F , fig.width = 8}

for(i in 1:2){
  
  assign(paste0("f",i),
         ggmcmc::ggs_traceplot(glm_mcmc_out2, 
                               family = names(glm_mcmc_out)[i])+
           theme(legend.position = "none") )
  
}

grid.arrange(f1,f2, nrow = 1)
```

```{r density_glm, fig.cap = "\\label{fig:density_glm}Density plot of entire chain (grey) and last 10\\% of the chain (green).", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8}

for(i in 1:2){
  
  assign(paste0("f",i),
         ggmcmc::ggs_compare_partial(glm_mcmc_out2, 
                                     family = names(glm_mcmc_out)[i]) +
           theme(legend.position = "none") )
  
}

grid.arrange(f1,f2, nrow=1)

```

Finally, we examine the values of all parameters through a caterpillar plot and compare the prior and posterior distributions.

```{r caterpillar_glm,   fig.cap = "\\label{fig:caterpillar_glm}Caterpillar plot of all parameters.", fig.align='center', fig.width = 8, fig.pos = 'h', strip.white=F}

ggmcmc::ggs_caterpillar(glm_mcmc_out2)
```

```{r priorpost_glm,   fig.cap = "\\label{fig:priorpost_glm}Prior (colour) and posterior (grey) distributions for all parameters.", fig.align='center', fig.pos = 'h', strip.white=F, fig.width = 8, fig.height=10}

plot_prior_post(glm_mcmc_out, 
                prior_col = snapalette::snapalette(snapal, 
                                                   3, 
                                                   type="discrete")[2], 
                "GLM")
```

\FloatBarrier 

## Assessing the fit

We now collect the estimates to produce the posterior predictive distribution of yellow fever reports.

```{r glm_predict,  fig.cap = "\\label{fig:predict_glm}Prediction of glm. The scale indicates the probability of  YF report over the observation period.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 5, fig.width=8}


Est_beta =  apply(glm_mcmc_out[,1:20], 2, median)

# read shapefiles in #
shp0 = readShapePoly(paste0("../../../shapefiles/gadm2/", "Africa_adm0.shp")) #if gadm2
shp1 = readShapePoly(paste0("../../../shapefiles/gadm2/", "Africa_adm1.shp"))

# adjust titles #
shp1$adm0_adm1 = paste(shp1$ISO, shp1$ID_1, sep="_")
shp1 = shp1[order(shp1$adm0_adm1),]

# read countries in #
Countries = read_csv(paste0("../../../Data/","Countries.csv"))

#sort age vector
ageVec = c(0:100)

# LOAD ENVIRONMENTAL DATA #
Env_Table_path = (paste0("../../../Data/","Environment/Africa_adm1_dat_2017.csv")) 
envdat = launch_env_dat(Env_Table_path, Countries$c34)


# FIT GLM #

#read in models
Models= readr::read_csv(paste0("../../../YellowFeverModelEstimation2017/","Models.csv" ) ) 

object_glm = fit_glm(dat = envdat$dat, 
                     depi = envdat$depi, 
                     Models$fm )  

beta0 = object_glm[[1]]
x = object_glm[[2]]
y = object_glm[[3]]

#########################################################################################################
plot_glm_map(shp0, shp1, Countries$c34, envdat$dat, Est_beta, x, 
                                   snapalette::snapalette(snapal, 1000, type="continuous"))

```
