---
title: "Running the seroprevalence estimation"
author: "Katy Gaythorpe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the seroprevalence estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This is a guide on how to run the estimation for the transmission models of yellow fever seroprevalence. It is quite specific to the data input types and format that we use. Hopefully, in future, this will be more general.

## Setup

```{r estimation_setup, eval=FALSE}

# load libraries #
library(dplyr)
library(readr)
library(reshape2)
library(abind)
library(mvtnorm)
library(truncdist)

library(YFburden)
library(YFestimation)

# load countries #
Countries = read_csv(paste0("../Data/","Countries.csv"))

# load serology data #
Serology = read.csv(paste0("../Data/","Serology/serology.csv"), 
                    stringsAsFactors = FALSE)
seroout = process_serology(Serology)

# load environmental data # 
# THIS MAY BECOME OBSOLETE #
Env_Table_path = paste0("../Data/","Environment/Africa_adm1_dat_2017.csv") 
envdat = launch_env_dat(Env_Table_path, Countries$c34)

# load population data #
all_res_pop_3d = get_pop_data_3d(path = "../Data/", 
                                 c_country = Countries$c34, 
                                 dat = envdat$dat) 

pop1 = all_res_pop_3d$pop1                                        
pop3d = all_res_pop_3d$pop3d                                      
P_tot_2d = all_res_pop_3d$P_tot_2d   
p_prop_3d = all_res_pop_3d$p_prop_3d      

#get names
dim_adm  = dimnames(pop3d)[[1]]
dim_year = as.numeric(dimnames(pop3d)[[2]])
dim_age  = dimnames(pop3d)[[3]]
dim_survey = seroout$sero_studies

# load vaccination data #
vaccdir = paste0("../Data/", "Vaccination/", "Outputs/", "adm1_old/")

latest_vaccine_csv = "vaccination_coverage_by_adm1_year_age_base_skew0.csv"   

vc2d = read.csv(paste0(vaccdir,latest_vaccine_csv), 
                stringsAsFactors = FALSE) 

names(vc2d)[names(vc2d)=="country"]= "adm0"       
names(vc2d)[names(vc2d)=="adm1"]= "adm0_adm1"    


for (colIndex in 3:ncol(vc2d)){  
  #before 1995, we have NA values for those aged >75
  vc2d[,colIndex] = ifelse(is.na(vc2d[,colIndex]),
                           vc2d[,colIndex-1], 
                           vc2d[,colIndex])
}
# restrict to lines in dat
vc2d = vc2d[vc2d[,"adm0_adm1"] %in% envdat$dat[,"adm0_adm1"],]

```

## Aggregate data to survey level and time

```{r aggregate}
list_aggregate_pop_vc = Make_aggregate_pop_vc_3d(pop1=pop1, 
                                                 vc2d=vc2d, 
                                                 sero_studies=seroout$sero_studies, 
                                                 adm1s=seroout$adm1s) 
pop_agg3d = list_aggregate_pop_vc$pop_agg3d 
vc_agg3d = list_aggregate_pop_vc$vc_agg3d 

#calculate aggregated incidence (same function as before)
inc_v3d_agg = calc_incidence_vac_general(vc_agg3d) 

#calculate aggregated moments (different fucntion before)
pop_moments_agg = calc_pop_moments_agg(pop_agg3d,
                                       seroout$t0_vac,
                                       dim_year,
                                       seroout$study_years) 

```

## R0 lookup table

This is not needed directly for the estimation so this section will be moved.

```{r R0_lookup}

if(!file.exists(paste0("../YellowFeverModelEstimation2017/",
                       "R0_lookup_table.Rdata") )){
  
  vc3d = transform_into_vc3d(vc2d,  adm="adm1")
  
  inc_v3d = calc_incidence_vac_general(vc3d)
  
  t0_vac_africa = calc_t0_vac_africa(vc3d)
  
  pop_moments_whole = calc_pop_moments(p_prop_3d, 
                                       t0_vac_africa,
                                       dim_adm,
                                       dim_year,
                                       dim_age)
  
  ptm = proc.time()
  create_R0_lookup(dim_adm, 
                   envdat$dat, 
                   dim_year,
                   dim_age,
                   p_prop_3d,
                   P_tot_2d,
                   inc_v3d,
                   pop_moments_whole,
                   pop_moments_agg, 
                   vac_eff_arg = 0.975)
  proc.time()-ptm
} else {
  load(paste0("../YellowFeverModelEstimation2017/",
              "R0_lookup_table.Rdata") )
}

```


## Set the initial parameters

```{r inital_param}

# additional foi to add to CAF survey #
foi_const_surv = c(0,1e-6,0,0,0,0,rep(0,seroout$no_sero_surveys-6))

# get population sizes at surveys #
out_p = create_pop_at_survey(pop_agg3d,dim_survey,dim_year) 
p_at_survey = out_p$p_at_survey_3d
P_tot_survey = out_p$P_tot_survey_2d

# load parameters from a previous run #
StartParamFoi=read.csv(paste0("../YellowFeverModelEstimation2017/",
                              "StartParam_","Foi",".csv"))

# itialising parameters for estimation #
parnames =  c("vac_eff",
              paste("Foi", seroout$sero_studies, sep = "_"),
              paste("R0", seroout$sero_studies, sep = "_"),
              paste("vc_factor_CMRs", sep = "_")
)
pars_ini = rep(NA,length(parnames))
names(pars_ini) = parnames

## filling the pars_in vector
# vaccine efficacy 
pars_ini[1] =log(0.83) 

# foi for each survey 
pars_ini[grep("Foi", parnames)] = StartParamFoi[(21+1) : (21+seroout$no_sero_surveys),1] 

# R0 for each survey KATY IS LOG TRANSFORMING THESE
#this is the max post prob values of R0 from trial run
pars_ini[grep("R0", parnames)] = c(0.1503953022, 0.3708065293, 0.3441911224, 0.5709173667, 0.1821695028, 
                                   0.1586931611, 0.0016357683, 0.0121969578, 0.0391997444, 0.0110269535, 
                                   0.0226958932, 0.0033852389, 0.0033071680, 0.0051427979, 0.0012126984, 
                                   0.0029272731, 0.0039877790, 0.0246799862, 0.0143955347, 0.0145107003, 
                                   0.0306652264, 0.0007623095, 0.0197133792, 0.0009735660, 0.0029589214, 
                                   0.0074393355, 0.0203410934, 0.0071625044, 0.0104076848, 0.0017355270, 
                                   0.0012662915, 0.0411910100, 0.0473851303, 0.0648086996, 0.0144261102, 
                                   0.0121657436, 0.0456392304, 0.0270631846, 0.2184448114, 0.1889724788) 

#vc.factor.CMRs 
pars_ini[length(pars_ini)]= -0.3184868 

# indices where vaccination affects serology #
varsin_nc = c(1, 17, 18, 19, 20)

# declare vector to identify different parameter types: vacc eff=1, Foi/R0=3, vc.factor.CMRs =4
parameter_type = c(1,rep(3,2*seroout$no_sero_surveys), 4) # THIS NOW INDEXES THE DIFFERENT PARAMETER TYPES

# initial model #
model_type = "R0" #"Foi" #

```

## Set pseudo prior distributions

```{r pseudopriors}

# load posteriors from previous runs #
FOI_posterior_distributions <-
  read.csv(paste0("Z:/MultiModelInference/", 
                  "Foi/posterior_distributions_norm.csv"),
    stringsAsFactors = FALSE)
FOI_posterior_distributions = cbind(FOI_posterior_distributions, model_type = "Foi")

R0_posterior_distributions <- 
  read.csv(paste0("Z:/MultiModelInference/", 
                  "R0/posterior_distributions_norm_tweak.csv"), 
    stringsAsFactors = FALSE)
R0_posterior_distributions = cbind(R0_posterior_distributions, model_type = "R0")

# bind #
posterior_distributions = rbind(FOI_posterior_distributions, 
                                R0_posterior_distributions)

#tweak posterior distributions 
scale_factor = setup_pseudoprior(pars_ini, 
                                 posterior_distributions, 
                                 "R0")

posterior_distributions$sd[posterior_distributions$model_type == "R0"] = 
  scale_factor * posterior_distributions$sd[posterior_distributions$model_type == "R0"]

```

## Product space MCMC

This is a transdimensional Markov chain Monte Carlo estimation technique specified for two models and sampled through Metropolis-within-Gibbs. See Gaythorpe et al. 2019 for further details.

```{r estimation}

#whether to ignore any surveys from likelihood calculation
ign = 2

#calculate the log probability of the Foi model for the intial parameters- 
#this will need adjusting so do not get too comfortable
prob_Foi = setup_modelprior(pars_ini,
                            seroout,
                            foi_const_surv,
                            vc_agg3d,
                            pop_agg3d,
                            pop_moments_agg,
                            dim_year,
                            dim_age,
                            p_at_survey,
                            P_tot_survey,
                            inc_v3d_agg, 
                            ign )

#create a directory to save the output in
name_dir = paste0("multi_model_MCMC_chain", "_", 
                  format(Sys.time(),"%Y%m%d"), 
                  "_", run_id)
dir.create(name_dir,  showWarnings = TRUE)

#specify number of iterations
Niter = 5e4

Sero_Gibbs_MCMC(pars_ini = pars_ini,
                seroout = seroout,
                foi_const_surv = foi_const_surv,
                vc_agg3d = vc_agg3d,
                pop_agg3d = pop_agg3d,
                pop_moments_agg=pop_moments_agg,
                dim_year = dim_year,
                dim_age = dim_age,
                p_at_survey = p_at_survey,
                P_tot_survey = P_tot_survey,
                inc_v3d_agg = inc_v3d_agg,
                model_type = model_type,
                parameter_type = parameter_type,
                prob_Foi,
                posterior_distributions = posterior_distributions,
                ign,
                name_dir = name_dir,
                Niter = Niter)

```
